Your APA6-Style Manuscript

### 2021-01-28 analysis/manuscript.Rmd
introduction

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
library(stringr)
```

@@

Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized-—for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.

>>>>

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
library(stringr)
```
# Introduction

The rise of open access (OA) to journal articles has added complexity to scholarly publishing, particularly concerning transparency of economic dimensions. Financial transparency in journal publishing has long been lacking since Big Deal subscription contracts between academic institutions and large academic publishers usually include confidentiality clauses that prohibit publicly sharing information about the scope and pricing of such agreements [@Frazier_2001; @Bergstrom_2014; @Larivi_re_2015]. Further, the distributed process of paying for accessing and publishing scholarly articles, which can involve several actors at the same time, including authors, research organizations, libraries, and research funders, adds to the opacity of the financial flows of scholarly publishing [@Lawson_2016].

The problem with this lack of transparency becomes more apparent in the case of hybrid OA. In this OA business model, individual articles can be made openly available for a fee, also known as article-processing charge (APC), while the journal as a whole remains subscription access. Hence, hybrid OA was originally introduced as a low-risk transitional model that allows journals to gradually convert to full OA and reduce subscription costs as the OA uptake increases [@Prosser_2003]. Many large subscription publishers have since incorporated this OA model, and while the uptake among authors increased only slowly during the first decade [@Bj_rk_2012], it has gradually gained popularity in recent years. This development was driven by the implementation of OA policies by research institutions and funders, the allocation of OA funds, and formal agreements between publishers and institutions that allow affiliated authors to publish OA free-of-charge. However, there have been concerns that the hybrid model allows for double-dipping—receiving two payments for one article, the APC and subscription fees [@Shieber_2009; @Prosser_2015]. While publishers assure they combine subscription fees and APCs to adjust pricing [@Mittermaier_2015], without transparency around the uptake of hybrid OA and revenue streams, such claims are impossible to evaluate.

With the recent introduction of transformative agreements, an evolving term for contracts that shift library spendings for subscriptions to OA [@Borrego_2020; @Hinchliffe_2019], the demand for publisher-provided data has even increased. If publishers would enable transparency about the OA uptake in hybrid journals and APC invoicing, academic libraries could improve their assessments and adjustments of contracts with publishers [@antelmann_2017]. Likewise, research organizations and funders could improve compliance monitoring with OA mandates, while ensuring that the different actors were not charged multiple times [@Larivi_re_2018]. However, as formal bibliometrics and financial studies have noted, the lack of publicly available and standardized data about hybrid OA publishing is considerable [@Laakso_2016; @Lawson_2015; @Pinfield_2016]

This current study investigates the extent to which publisher-provided data can be utilized to assess hybrid OA. The study takes Elsevier, the largest scholarly journal publisher in terms of article volume [@Larivi_re_2015], as the starting point for an empirical analysis of hybrid OA uptake and invoicing. Our approach is based on openly available data sources and brings together scholarly data from Crossref and machine-readable APC invoicing data from Elsevier OA full-texts, which enables us to analyze the development of Elsevier’s hybrid journal portfolio from 2015-19, and to compare it with that of Elsevier’s full and delayed OA program. Further, we investigate whether hybrid publishing fees were waived or charged to the authors or academic institutions or research funders with Elsevier OA agreements. In particular, we identify academic consortia and research funders that were directly invoiced by Elsevier, and compare their hybrid OA rates by license and subject. As such, our findings have implications for research and OA policy and highlight the potential of publisher-provided data for large-scale, instantaneous analysis of hybrid OA uptake and invoicing.

@@

Since there is no standardized way to measure the uptake of hybrid OA, estimates can vary greatly depending on how uptake is operationalized--for example, through the number of hybrid journals with at least one OA article, the average number of OA articles per hybrid journal, or the share of OA articles compared to subscription articles.
